Policy optimization for dynamic spatiotemporal systems \hspace{1em} \insertframenumber

### 2014-08-05 pot/LaTeX/jsm2014.tex
mapping, slide updates
Edit to mapping file. Pull out different information. Slide updates

\begin{frame}{Standard non-spatial analysis}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Block maxima:
    \begin{itemize}
      \item Uses yearly maxima
      \item Discards many observations
      \item Models are fit using the generalized extreme value distribution
    \end{itemize}
    \item Generalized extreme value distribution (GEV):
    \begin{align*}
      \Pr(Y_j < y) = G_j(y) = \exp \left\{ -\left[ \left(1 + \xi_j \frac{y - \mu_j}{\sigma_j}\right)_+^{-1/\xi_j} \right] \right\}
    \end{align*}
  \end{itemize}
\end{frame}

@@

@@

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item \alert{Conditioned} on $z_t$ and $\sigma^2_t$, $Y_t(\bs)$ is Gaussian
    \item Can use standard geostatistical methods to fit this model.
    \item Predictions can be made through kriging.
    \item \alert{Marginalizing} over $z_t$ and $\sigma^2_t$ (via MCMC),
    \begin{align*}
      Y_t(\bs) \sim \text{skew-t}(\mu, \Sigma^*, \alpha, \text{df}=2a)
    \end{align*}
    where
    \begin{itemize}
    	\item $\mu$ is the location
	\item $a$, $b$ are the IG parameters for $\sigma^2_t$
	\item $\Sigma^* = \frac{ b }{ a } \Sigma$ is a scale matrix, and $\Sigma$ is a \Matern covariance matrix
	\item $\alpha \in \calR$ controls the skewness
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Long-range dependence}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item The $\chi$ coefficient is a measure of extremal spatial correlation
    \begin{align*}
      \chi(\bh) = \Pr(Y_t(\bs) > c \mid Y_t(\bs + \bh) > c)
    \end{align*}
    \item This value shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$ (Padoan, 2011)
    \item Deal with this through a daily random partition.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Random daily partition}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Daily random partition allows $z_t$ and $\sigma^2_t$ to vary by site.
    \begin{align*}
      Y_t(\bs) = X_t(\bs) \beta + \alpha z_t(\bs) + \sigma(\bs) v_t(\bs)
    \end{align*}
    \item Consider a set of daily knots $\{w_{t1}, \ldots, w_{tK}\}$ that define a daily partition
    $P_{t1}, \ldots, P_{tK}$ such that
    \begin{align*}
      P_{tk} = \{s : k = \argmin_\ell|| \bs - w_{t\ell}|| \}
    \end{align*}
    \item For $\bs \in P_{tk}$
    \begin{align*}
      z_t(\bs) &= z_{tk}\\
      \sigma^2_t(\bs) &= \sigma^2_{tk}
    \end{align*}
    \item Within each partition $Y_t(\bs)$ has the same MVT distribution as before.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Future work}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Comparison with extreme value analysis methods
    \item Including time in the model
    \begin{itemize}
      \item AR(1): $Y_t(\bs) = X_t(\bs) \beta + \phi Y_{t-1}(\bs) + \alpha z_t(\bs) + v_t(\bs)$
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{References}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Huser, R. and Davison, A. C. (2014) Space-time modelling of extreme events. {\it Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, {\bf 76}, 439--461.
    \item Padoan, S. A. (2011) Multivariate extreme models based on underlying skew-$t$ and skew-normal distributions. {\it Journal of Multivariate Analysis}, {\bf 102}, 977--991.
    \item Zhang, H. and El-Shaarawi, A. (2010) On spatial skew-Gaussian processes and applications. {\it Environmetrics}, {\bf 21}, 33--47.
  \end{itemize}
\end{frame}

>>>>

\begin{frame}{Standard non-spatial analysis}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Block maxima:
    \begin{itemize}
      \item Uses yearly maxima
      \item Discards many observations
      \item Models are fit using the generalized extreme value distribution
    \end{itemize}
    \item Generalized extreme value distribution (GEV):
    \begin{align*}
      \Pr(Y_j < y) = G_j(y) = \exp \left\{ -\left[ \left(1 + \xi_j \frac{y - \mu_j}{\sigma_j}\right)_+^{-1/\xi_j} \right] \right\}
    \end{align*}
    \item For a spatial analysis, max-stable processes give an appropriate limiting distribution.
  \end{itemize}
\end{frame}

@@

\begin{frame}{$\chi$ coefficient}
  \begin{itemize} \setlength{\itemsep}{0.5em}
   \item The $\chi$ coefficient is a measure of extremal correlation
   \item Specifically, we focus on $\chi(c, \bh)$ for the upper tail given by
    \begin{align*}
      \chi(c, \bh) = \Pr(Y(\bs) > c \mid Y(\bs + \bh) > c)
    \end{align*}
    \item If $\lim_{c \rightarrow \infty} \chi(c) = 0$, then high extreme values are not correlated.
  \end{itemize}
\end{frame}

\begin{frame}{Gaussian spatial model}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item In geostatistical models, observations, $Y(\bs)$ are often modeled using a Gaussian spatial model with a likelihood of
      \begin{align*}
    	\hspace{-2em}
    	f(y(\bs)) = \frac{ 1 }{ (2 \pi)^{n/2} } | \Sigma |^{-1/2} \exp \left[ -\frac{ 1 }{ 2 } (Y(\bs) - \mu(\bs) )^T \Sigma^{-1} (Y(\bs) - \mu(\bs)) \right]
    \end{align*}
    where
    \begin{itemize}
    	\item $\mu(\bs)$ is a mean vector
	\item $\Sigma$ is a covariance matrix
    \end{itemize}
    \item Model properties:
    \begin{itemize}
      \item Nice computing properties
    \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(c) = 0$ regardless of the strength of the correlation in the bulk of the distribution.
    \item Tail is not flexible
    \end{itemize}
    \end{itemize}
\end{frame}

@@

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
   \item \alert{Conditioned} on $z_t$ and $\sigma^2_t$, $Y_t(\bs)$ is a Gaussian spatial model
    \item Can use standard geostatistical methods to fit this model.
    \item Predictions can be made through kriging.
    \item \alert{Marginalizing} over $z_t$ and $\sigma^2_t$ (via MCMC),
    \begin{align*}
      Y_t(\bs) \sim \text{skew-t}(\mu, \Sigma^*, \alpha, \text{df}=2a)
    \end{align*}
    where
    \begin{itemize}
    	\item $\mu$ is the location
	\item $a$, $b$ are the IG parameters for $\sigma^2_t$
	\item $\Sigma^* = \frac{ b }{ a } \Sigma$ is a scale matrix, and $\Sigma$ is a \Matern covariance matrix
	\item $\alpha \in \calR$ controls the skewness
    \end{itemize}

  \end{itemize}
\end{frame}

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Model properties
    \begin{itemize}
    	\item Has flexible tail controlled by $\alpha$ and degrees of freedom
    	\item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(c) > 0$ Padoan, 2011)
    	\item Computation that is on the order of Gaussian computation
    \end{itemize}
    \item For this distribution, $\chi(c, \bh)$ shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$ 
    \item Deal with this through a daily random partition.
  \end{itemize}
\end{frame}

\begin{frame}{Daily random partition}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Daily random partition allows $z_t$ and $\sigma^2_t$ to vary by site.
    \begin{align*}
      Y_t(\bs) = X_t(\bs) \beta + \alpha z_t(\bs) + \sigma(\bs) v_t(\bs)
    \end{align*}
    \item Consider a set of daily knots $\{w_{t1}, \ldots, w_{tK}\}$ that define a daily partition
    $P_{t1}, \ldots, P_{tK}$ such that
    \begin{align*}
      P_{tk} = \{s : k = \argmin_\ell|| \bs - w_{t\ell}|| \}
    \end{align*}
    \item For $\bs \in P_{tk}$
    \begin{align*}
      z_t(\bs) &= z_{tk}\\
      \sigma^2_t(\bs) &= \sigma^2_{tk}
    \end{align*}
    \item Within each partition $Y_t(\bs)$ has the same MVT distribution as before.
  \end{itemize}
\end{frame}

@@

@@

\begin{frame}{Future work}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Comparison with extreme value analysis methods
    \item Reporting of compliance with EPA ozone standards
    \item Spatial mapping of high quantiles
    \item Including time in the model
    \begin{itemize}
      \item AR(1): $Y_t(\bs) = X_t(\bs) \beta + \phi Y_{t-1}(\bs) + \alpha z_t(\bs) + v_t(\bs)$
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{References}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Demarta, S. and McNeil, A. J. (2007) The $t$ copula and related copulas. {\it International Statistical Review}, {\bf 73}, 111--129.
    \item Huser, R. and Davison, A. C. (2014) Space-time modelling of extreme events. {\it Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, {\bf 76}, 439--461.
    \item Padoan, S. A. (2011) Multivariate extreme models based on underlying skew-$t$ and skew-normal distributions. {\it Journal of Multivariate Analysis}, {\bf 102}, 977--991.
    \item Zhang, H. and El-Shaarawi, A. (2010) On spatial skew-Gaussian processes and applications. {\it Environmetrics}, {\bf 21}, 33--47.
  \end{itemize}
\end{frame}
